---
title: "EDF 5401 Homework Assignment 1"
format: pdf
editor_options: 
  markdown: 
    wrap: 72
---df
---

(total points: 90)

Please read all questions carefully. If you are working with a group to
complete this assignment, all members of your group must sign the
statement below:

**All members of this homework team contributed equitably to the
assignment.**

**Member 1:**

**Member 2:**

**Member 3:**

```{r libraries}
library(DescTools)
library(tidyverse)
```

This assignment is based on the data file "jobsat.csv" (or "jobsat.sav"
for SPSS users). This file contains 9 variables from 200 employees from
a certain type of company, plus a variable "ID". The meaning of each
variable is as follow:

-   Gender: 1=female; 2=male

-   Age: the age of employee

-   Physical environment: a scale score indicating how employees like
    the current physical environment in their companies; scale scores
    range from 1 to 30 with higher scores indicating positive feeling
    about the environment.

-   Performance: a rating on employee's previous year's job performance
    using a scale of 1 to 50 with higher rating suggesting better
    performance.

-   Previous year salary: previous year's salary in thousands of dollars

-   Current salary: current annual salary in thousands of dollars

-   Stress: employee's perceived stress for their job, higher score
    suggesting higher level of stress.

-   Job satisfaction: satisfaction for the current job; scale scores
    range from 1 to 50 with higher scores indicating higher level of
    satisfaction.

-   Rating (rating): supervisor's rating on this year's performance as
    measured by 0 (unsatisfactory) or 1 (satisfactory)

```{r loadData}
jobsat <- read_csv("jobsat.csv")
```

In this homework assignment, you only need to consider two variables:
job satisfaction and current salary (plus the ID number). You try to
describe and understand the relationship between job satisfaction and
current salary. Commonly speaking, employee tends to feel satisfied with
their current job if they perceived they are well-paid. You man use R,
SPSS, or any other stat package you are familiar with.

1.  (5 pts.) Suppose you want to predict one of the two variables from
    the other. Based on the description above and your understanding of
    the relationship between job satisfaction and current salary, which
    variable do you think is more appropriate to be conceptualized as
    outcome variable, and which as predictor?

2.  (5 pts.) Use a statistics package to compute descriptive statistics
    (e.g., mean, median, standard deviation, skewness and kurtosis).
    Make two histograms or Q-Q plots, one for job satisfaction and the
    other for current salary. Do these two variables seem normally
    distributed?

```{r Q2}

```

3.  (5 pts.) Use a statistics package to make a scatterplot of job
    satisfaction versus current salary. How would you characterize the
    relationship between job satisfaction and current salary? Be sure to
    describe the relationship in terms of three aspects: direction,
    strength, and shape.

```{r Q3}

```

4.  (5 pts.) Check the scatterplot. Do any point(s) appear to be
    outlier(s)? What evidence supports your choice(s)? (Hint: making
    boxplots might be useful).

```{r Q4}

```

5.  (5 pts.) Obtain the bivariate correlation between job satisfaction
    and current salary. Does the correlation you have obtained support
    your description in Question 3?

```{r Q5}

```

6.  You have defined outcome variable and predictor in the Question 1.
    Conduct simple regression that predicts the outcome variable from
    the predictor using this sample of data. Save residuals from the
    model, as well as dfbetas. Then answer the following questions:

```{r Q6}

```

6.1 (5 pts.) Write equations for this regression model at the population
level (two equations: one for observed scores and the other for the
regression line). Define each element in the equation.

> In Rmarkdown, you can make an equation by enclosing it in a pair of
> `$`s. Displayed equations have double dollar signs. Within an
> equation, `_` is a subscript and `^` is a superscript. Greek letters
> are made by using a backslash (`\`) followed by the name of the
> letter. For example, `\epsilon` becomes $\epsilon$.

Observed Scores:

$$
  Y_i =
$$ Regression Line:

$$
  Y =
$$

(6.2) (5 pts.) Check "Coefficients" table. What are the values of $b_0$
and $b_1$ the unstandardized regression equations for this model (two
equations: one for observed scores and the other for the regression
line).

(6.3) (5 pts.) Interpret the value of $b_0$ and $b_1$.

(6.4) (5 pts.) Obtain the value of MSE and SEE. What do these two values
tell you? Comparing these two values to the variance and standard
deviation of the outcome variable, does the predictor seem to help in
predicting the outcome variable?

(6.5) (5 pts.) Check residuals and dfbetas, which case(s) seem to be
outlier(s)? Support your answer with appropriate indices. Explain in
words the meaning of the largest dfbeta value for $b_1$.

7.  Exclude the outlier(s) you identified in the Question 6.5 and re-run
    the simple regression. Save residuals and dfbetas from the model,
    and make a plot with standardized residuals vs. predicted $Y$
    values. Then answer the following questions:

```{r Q7}

```

(7.1) (5 pts.) What are the value of slope ($b_1$) and its standard
error ($s_{b_1}$)? Use these two values to conduct the test of H0:
$\beta_1= 0$ at $\alpha = .05$. What is your $t$ value? $p$ value? What
is your conclusion about the effect of the predictor on the outcome
variable?

(7.2) (5 pts.) What's the null hypothesis of the $F$ test? Using values
from the ANOVA table, compute the $F$ value. What is your $p$ value?
Suppose $\alpha = .05$, what is your statistical conclusion for this
hypothesis testing?

(7.3) (5 pts.) Square the $t$ value for the slope from Question 7.1, and
compare it to the $F$ value from Question 7.2. Explain your findings. Do
the \$F%-test and $t$\*-test lead to the same statistical conclusion?

(7.4) (5 pts.) Compute the 95% confidence interval (C.I.) around the
estimated slope ($b_1$). What does this interval tell you about the
population value of slope? What does this interval suggest about the
relationship between job satisfaction and current salary? Compare this
information to that given by the t test in Question 7.1, does this C.I.
lead to the same statistical conclusion as the $t$-test?

(7.5) (5 pts.) Describe the relationship between $R^2$ and the
quantities in the ANOVA table. Interpret the meaning of this value.

(7.6) (5 pts.) Do you think this predictor (current salary) is important
in predicting the outcome variable (job satisfaction)? Why do you think
it is important (or not important)?

(7.7) (5 pts.) Check the scatterplot between standardized residuals and
predicted $Y$ values (SPSS default) or the scatterplot between square
root of the absolute value of the residuals (R default). Does the
assumption of homogeneity of error variance (homoscedasticity) seem to
be met for this regression analysis?

(7.8) (5pts.) Does the fitted model excluding the outliers seem to be
similar to or different from the model based on all 200 cases? Comment
on any differences you see between these two models.
