---
Title: 'Homework 3:  Logistic Regression'
output: word_document
---

```{r}
library(tidyverse)
library(DiagrammeR)
```

# A Quick Introduction to Knowledge Tracing

The knowledge tracing model [(Corbet and Anderson, 1995)](CorbettAnderson1995.pdf) posits
that there is a latent skill, whose value at Time $t$ is $L_t$.  Each time the student
practices the skill there is a chance that the student learns the skill ($L_{t+1} =1$).
The diagram below shows this struction, where $L_t$ are the latent skills, and $Y_t$ are
the observed correct/incorrect scores for the items.


```{r}
DiagrammeR::grViz('
digraph KT {
  rankdir="BT"
{rank="max"
  L1 -> L2 -> L3 -> L4 -> L5 -> L6 -> L7 -> L8
}
{rank="min"
  Y1; Y2; Y3; Y4; Y5; Y6; Y7; Y8
}

L1 -> Y1
L2 -> Y2
L3 -> Y3
L4 -> Y4
L5 -> Y5
L6 -> Y6
L7 -> Y7
L8 -> Y8

}')
```

Following [Van de Sande, 2013](VandeSande2013.pdf), this model has four parameters.

- $P(L_0)$ -- Probability student has the skill at Time 0 (before the first exercise).
- $P(T)$ -- Probability that a student who is a non-master makes the transition to mastery 
at a practice occasion
- $P(G)$ -- Probability that a non-master gets the item correct (guessing)
- $P(S)$ -- Probabiity that a master gets the item incorrect (slipping)

After some algebra, it can be seen that the students probability of success at Time $t$ can
be written as 

$$ P(Y_t=0) = \textrm{logit}^{-1}( \alpha + \beta t) \ , $$

where $\beta = \log(1-P(T))$ (and $\alpha$ is a complex function of all of the parameters.
Thus we can fit this model by using logistic regression.

# The Data

The data in [`homework.csv`](homework.csv) is simulated data from a homework system.  There are 100 students and there are four variables.

- `SID` the student ID
- `group` a demographic group, which takes the values `Focal` and `Reference`. (Focal and 
reference are names used in differential item functioning analysis, usually the "focal" 
group is one which has some history of discrimination, and the reference group is a higher
status comparison group.)
- `attempt` the trial number (or time in the above discussion)
- `success` A logical value indicating whether the student was success at that trial or not.


```{r}
homework <- read.csv("homework.csv")
homework$group <- as.factor(homework$group)
summary(homework)
```

# The assignment

## 1. Learning curve

Calculate the fraction of successful trials for each time point (1 through 8).
Plot the trial number versus the success probability.  This is known as the _learning 
curve_.

## 2. Plot the data with an estimated curve.

Note:  you will want to jitter the values so they can be seen, and also add a lowess curve.

## 3. Fit a logistic regression

Use the model `success ~ attempt`
Write up a summary of the results.

## 4. Interpret the slope

Explain in words what the slope means.

## 5. Add color to the plot according to the `group`

Describe the difference if any between the groups.

## 6. Pick the right model with the group

Fit both the model `success ~ attempt + group` (no interaction with slope) and
`success ~ attempt * group` (adds an interaction) and compare them to the no group model 
`success ~ attempt` which you fit before.  Which model do you like better and 
why?

## 7. Describe the model you picked in words and equations.

## 8. Make a graph of the final model

Hint:  Predict the success rate for people in both the reference and focal groups at 
attempts 1 through 8 and plot that.


