---
title: "Midterm Blueprint"
format: docx
---

```{r}
library(tidyverse)
library(DescTools)
library(plotly)
```

## Generic Questions

Given a description of a data collection, and a bunch of R output.

1.  Are there problems with the sample? Non-representative (does it match the target population)? Missing data?

2.  Is the distribution the same for everybody in the group? Are there clusters? Serial (time) dependencies?

3.  What at the distributions of X and Y like? High skeweness? High kurtosis? Outliers?

4.  Is the relationship mostly linear? Will transforming X or Y make it more linear?

5.  Is the variance roughly the same for all values of $\hat Y$? Will transforming $Y$ help?

6.  Are the residuals roughly normal? Are there any outliers? Will transforming $Y$ help?

7.  Are there any differences by subgroup?

8.  Is there enough evidence in the sample to conclude that $X$ and $Y$ are related? How strong is the relationship?

9.  What is the equation of the regression line?

10. Sensitivity to outliers? Leverage points?

11. What can we conclude about the relationship between $X$ and $Y$?

# Part 1

Here is the analysis of ACED post_test scores by EAP scores.

## Exploratory Analysis

### Read the data

```{r loadACEDnoNA}
ACEDextract <- read_csv("../CaseStudies/ACED_extract1.csv",na="-999")
ACEDextract$Session <- factor(ACEDextract$Session)
ACEDextract$Cond_code <- factor(ACEDextract$Cond_code)
ACEDextract$Sequencing <- factor(ACEDextract$Sequencing)
ACEDextract$Feedback <- factor(ACEDextract$Feedback)
ACEDextract$Gender <- factor(ACEDextract$Gender)
ACEDextract$Race <- factor(ACEDextract$Race,1:8)
ACEDextract$Level_Code <- factor(ACEDextract$Level_Code)
```

### Marginal Summaries

```{r EAP}
Desc(ACEDextract$EAP.sgp)
```

```{r EAPnoOut}
Desc(ACEDextract$EAP.sgp[ACED_extract1$EAP.sgp<.33])
```

```{r post}
Desc(ACEDextract$post_scaled)
```

```{r outlierEAP}
select(ACEDextract,all_of(c("SubjID","Cond_code","Level_Code",
                            "post_scaled","EAP.sgp"))) %>%
  filter(post_scaled > 82)
```

### Scatterplots

```{r scatterplots0}
#| warning: false
#| message: false
ggplot(ACEDextract,aes(x=EAP.sgp,y=post_scaled)) + geom_point() +
  geom_smooth()
```

By study condition

```{r scatterplotsCC}
#| warning: false
#| message: false
ggplot(ACEDextract,aes(x=EAP.sgp,y=post_scaled,color=Cond_code)) + geom_point() +
  geom_smooth()
```

By class level

```{r scatterplotsLC}
#| warning: false
#| message: false
ggplot(ACEDextract,aes(x=EAP.sgp,y=post_scaled,color=Level_Code)) + geom_point() +
  geom_smooth(method="lm")
```

By class

```{r scatterplotsSession}
#| warning: false
#| message: false
ggplot(ACEDextract,aes(x=EAP.sgp,y=post_scaled,color=Session)) + geom_point() +
  geom_smooth(method="lm")
```

By Gender

```{r scatterplotsGender}
#| warning: false
#| message: false
ggplot(ACEDextract,aes(x=EAP.sgp,y=post_scaled,color=Gender)) + geom_point() +
  geom_smooth()
```

### First Regression

```{r ACEDregression}
aced_lm <- lm(post_scaled ~ EAP.sgp, data=ACEDextract,na.action=na.exclude)
summary(aced_lm)
```

```{r ACEDregression1}
aced_lm1 <- lm(post_scaled ~ EAP.sgp+pre_scaled, data=ACEDextract,na.action=na.exclude)
summary(aced_lm1)
```

```{r ACEDdiag}
oldpar <- par(mfrow=c(2,2))
plot(aced_lm)
par(oldpar)
```

### ELL Students

The accommodation given to English Language Learner (ELL) students was that their teachers could translated any words they were having problems with. However, the study team noted that these translations often became instruction.

Run without ELL students.

```{r ACEDnoELL}
aced_lm_noell <- lm(post_scaled ~ EAP.sgp, data=ACEDextract,
                    subset=ACEDextract$Level_Code!="ELL")
summary(aced_lm_noell)
```

```{r ACEDnoELLdiag}
oldpar <- par(mfrow=c(2,2))
plot(aced_lm_noell)
par(oldpar)
```

## Questions

1.  All the data were taken from a single suburban middle school in New Jersey. Does that present a problem for generalizing to other middle schools?

2.  The students were randomly assigned to study conditions, and no Control student has an `EAP.sgp` score (as they didn't use ACED). Does this introduce a bias?

3.  Is the distribution the same for everybody in the group? Are there differences by study condition (`Cond_code`)? class level (`Level_Code`)? class (`Session`)?

4.  What at the distributions of $X$ (`EAP.sgp`) and $Y$ (`post_scaled`) like? High skeweness? High kurtosis? Outliers?

5.  Is the relationship mostly linear? Will transforming $X$ or $Y$ make it more linear?

6.  Is the variance roughly the same for all values of $\hat Y$? Will transforming $Y$ help?

7.  Are the residuals roughly normal? Are there any outliers? Will transforming $Y$ help?

8.  Is there enough evidence in the sample to conclude that $X$ and $Y$ are related? How strong is the relationship?

9.  What is the equation of the regression line?

10. The ELL students had slightly different administration conditions. How sensitive are the conclusions to those students?

11. What can we conclude about the relationship between $X$ and $Y$? In particular, both `EAP.sgp` and `post_scaled` are measures of the students ability to solve geometric sequence problems? Are they measuring something similiar?

# Part 2

The data were originally published in the *Albaquerque Tribune* in 1986-11-07. It consists of data from 1985, from the 50 states plus the District of Columbia[^1]. It consists of the "average" (which average is unspecified, presumably the mean) teacher salary, and the per pupil spending on education in each space.

[^1]: Taxation without representation.

The data originally come from [https://lib.stat.cmu.edu/StatDat/Datafiles/teacherpaydat.html](https://web.archive.org/web/20110104102707/http://lib.stat.cmu.edu/DASL/Datafiles/teacherpaydat.html) (original link has gone, so this sends you to the Internet Wayback Machine). From this site:

1.  PAY (`Pay`): Average public school teacher annual salary (\$)
2.  SPEND (`Spend`): Spending on public schools per pupil \[sic\] (\$)
3.  AREA (`Region`): Region 1 Northeast and North Central, 2 South, 3 West

```{r loadTeacherPay}
teacherPay <- read_delim("TeacherPay.txt",trim_ws = TRUE)
teacherPay$Region <- factor(teacherPay$Region,1:3,
                            c("Northeast/North Central","South","West"))
summary(teacherPay)
```

## Exploratory Analysis

### Unidimensional

```{r pay}
Desc(teacherPay$Pay)
```

```{r spending}
Desc(teacherPay$Spend)
```

```{r outlier}
filter(teacherPay, Pay > 40000)
```

### Exploratory By Groups

```{r groups}
teacherPay %>% group_by(Region) %>%
  summarise(mPay = round(mean(Pay)), sdPay = round(sd(Pay)), 
            skPay=round(Skew(Pay),2),
            mSpend = round(mean(Spend)), sdSpend=round(mean(Spend)),
            skSpend=round(Skew(Pay),2)) 
```

```{r histByRegion}
#| warning: false
ggplot(teacherPay,aes(x=Pay)) + geom_histogram() +
  facet_grid(rows=teacherPay$Region)
```

```{r histByRegion2}
ggplot(teacherPay,aes(x=Spend)) + geom_histogram() +
  facet_grid(rows=teacherPay$Region)
```

### Scatterplots

```{r scatterplots}
#| warning: false
ggplot(teacherPay,aes(x=Spend,y=Pay,label=State)) + geom_point() +
  geom_smooth() + 
  geom_text(data=teacherPay[teacherPay$Pay>30000,],vjust=-1) +
  geom_text(data=teacherPay[teacherPay$Pay<20000,],vjust=1)
```

```{r scatterplotsC}
ggplot(teacherPay,aes(x=Spend,y=Pay,label=State,color=Region)) +
  geom_point() +
  geom_smooth()
```

### First Regression

```{r regression}
tp_lm <- lm(Pay ~ Spend, data=teacherPay)
summary(tp_lm)
```

### Diagnostic Plots

```{r lmPlot}
par(mfrow=c(2,2))
plot(tp_lm,labels.id=teacherPay$State)
par(mfrow=c(1,1))
```

### Log-log Model

Sometimes money variables are better on the log scale. Is this the case here?

```{r scatterplot}
#| warning: false
ggplot(teacherPay,aes(x=Spend,y=Pay)) + 
  scale_x_log10() + scale_y_log10() + 
  geom_point() + geom_smooth()
```

```{r logloglm}
tp_lllm <- lm(log(Pay) ~ log(Spend), data=teacherPay)
summary(tp_lllm)
```

```{r loglogdiag}
oldpar <- par(mfrow=c(2,2))
plot(tp_lllm,labels.id = teacherPay$State)
par(oldpar)
```

### Leverage points.

```{r dfbetas}
tp_dfb <- dfbetas(tp_lm)
summary(tp_dfb)
```

```{r highVals1}
tp_dfb <- data.frame(dfbetas(tp_lm))
tp_dfb$State=teacherPay$State
filter(tp_dfb,abs(Spend)>2/sqrt(nrow(tp_dfb)))
```

Save the row numbers for sensitivity analysis.

```{r highLev}
highLev <- which(abs(tp_dfb$Spend)>.3)
highLev
```

### Sensitivity Analysis 1, no Alaska

```{r lmnoAK}
tp_lm_noAK <- lm(Pay ~ Spend, data=teacherPay, 
                 subset=teacherPay$State != "AK")
summary(tp_lm_noAK)
```

```{r diagnoAK}
oldpar <- par(mfrow=c(2,2))
plot(tp_lm_noAK,labels.id = teacherPay$State)
par(oldpar)
```

### Take out all four high leverage points

R trick: using a negative index will select everything but selected values.

```{r lmLowLev}
tp_lm_lowLev <- lm(Pay ~ Spend, data=teacherPay, 
                   subset = -highLev)
summary(tp_lm_lowLev)
```

```{r diaglowlev}
oldpar <- par(mfrow=c(2,2))
plot(tp_lm_lowLev,labels.id = teacherPay$State)
par(oldpar)
```

### Table of Results

```{r extractor}
extractor <- function(lm_mod) {
   c(coef(lm_mod),df=summary(lm_mod)$df[2], 
            r.squared=summary(lm_mod)$r.squared,
            sigma=summary(lm_mod)$sigma)
}
stattab <- rbind(
  All=extractor(tp_lm),
  noAK=extractor(tp_lm_noAK),
  lowLev=extractor(tp_lm_lowLev)
)
stattab <- data.frame(stattab)
names(stattab)[1] <- "Intercept"
stattab$Intercept <- round(stattab$Intercept)
stattab$Spend <- round(stattab$Spend,2)
stattab$r.squared <- round(stattab$r.squared,2)
stattab$sigma <- round(stattab$sigma)
stattab
```

## Questions

1.  The data are from 1986. How much will that tell us about 2023?

2.  Are there any clusters in the data with different distributions?

3.  What at the distributions of $X$ (`Spend`) and $Y$ (`Pay`) like? High skeweness? High kurtosis? Outliers?

4.  Is the relationship mostly linear? Frequently with money, log transformations help. Which model is better `Pay ~ Spend` or `log(Pay) ~ log(Spend)`

5.  Is the variance roughly the same for all values of $\hat Y$? Will transforming $Y$ help?

6.  Are the residuals roughly normal? Are there any outliers? Will transforming $Y$ help?

7.  Are there any differences by region?

8.  Is there enough evidence in the sample to conclude that $X$ and $Y$ are related? How strong is the relationship?

9.  What is the equation of the regression line?

10. Which states are outliers in the regression?

11. Which states are high leverage points?

12. What is the sensitivity of the results to the high leverage values? Is it of concern?

13. What can we conclude about the relationship between $X$ and $Y$? In particular, does increasing spending cause an increase in pay? Increasing pay cause an increase in spending? Or is there some common factor.
