---
title: "Attendance"
author: "Russell Almond"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(GGally)
```

# Part 1 -- Multiple Regression

## Crime data set.  

These data come from <http://www.statsci.org/data/general/uscrime.html>.

They show various information about 47 states from 1960.

Here are the variable descriptions:

M
: percentage of males aged 14–24 in total state population

So
: indicator variable for a southern state

Ed
: mean years of schooling of the population aged 25 years or over

Po1
: per capita expenditure on police protection in 1960

Po2
: per capita expenditure on police protection in 1959

LF
: labour force participation rate of civilian urban males in the age-group 14-24

M.F
: number of males per 100 females

Pop
: state population in 1960 in hundred thousands

NW
: percentage of nonwhites in the population

U1
: unemployment rate of urban males 14–24

U2
: unemployment rate of urban males 35–39

Wealth
: median value of transferable assets or family income

Ineq
: income inequality---percentage of families earning below half the median income

Prob
: probability of imprisonment--ratio of number of commitments to number of offenses

Time
: average time in months served by offenders in state prisons before their first release

Crime
: crime rate---number of offenses per 100,000 population in 1960

The goal is to find what factors predict a high crime rate, that is predict crim.

## Read the data

```{r}
UScrime <- read.delim("http://www.statsci.org/data/general/uscrime.txt")
summary(UScrime)
```



## Scatterplot matrixes

We will do this a couple at a time so we can see the scatterplots.

```{r}
#| message: false
mutate(UScrime,Southern=as.logical(So)) |>
  ggpairs(columns=c("M","M.F","Southern","Ed","Crime"))
```
```{r}
#| message: false
mutate(UScrime,Southern=as.logical(So)) |>
  ggpairs(columns=c("Po1","Po2","LF","Pop","Crime"))
```
```{r}
#| message: false
mutate(UScrime,Southern=as.logical(So)) |>
  ggpairs(columns=c("NW","U1","U2","Wealth","Crime"))
```

```{r}
#| message: false
mutate(UScrime,Southern=as.logical(So)) |>
  ggpairs(columns=c("Wealth","Ineq","Prob","Time","Crime"))
```

## Scatterplot with Policing

```{r}
ggplot(UScrime,aes(x=Po1,y=Crime)) + geom_point() + 
  geom_smooth()
```
## Scatterplot with Policing

```{r}
ggplot(UScrime,aes(x=Po1,y=Crime,color=as.logical(So))) + geom_point() + 
  geom_smooth()
```

## Model with just policing

```{r}
lmPo1 <- lm(Crime~Po1,data=UScrime)
summary(lmPo1)
```

### Diagnostic plots

Using the `which` argument allow us to look at the plots one at a time.  (The default
uses plots 1, 2, 3 and 5).

```{r}
plot(lmPo1,which=1)
```
```{r}
plot(lmPo1,which=2)
```

```{r}
plot(lmPo1,which=3)
```

```{r}
plot(lmPo1,which=5)
```

## Data using log crime rate as Y

```{r}
lmlPo1 <- lm(log(Crime)~Po1,data=UScrime)
summary(lmlPo1)
```

```{r}
plot(lmlPo1,which=1)
```
```{r}
plot(lmlPo1,which=2)
```
```{r}
plot(lmlPo1,which=3)
```
```{r}
plot(lmlPo1,which=5)
```

## Include Policing in both years

```{r}
lmlPo12 <- lm(log(Crime) ~ Po1 + Po2,data=UScrime)
summary(lmlPo12)
```

## Model Search

```{r}
lmlstep <- step(lmlPo1,list(lower=log(Crime)~1,
                            upper=log(Crime)~M+So+Ed+Po1+Po2+LF+M.F+Pop+
                              NW+U1+U2+Wealth+Ineq+Prob+Time),
                            trace=2)
```

## Summary of final model

```{r}
summary(lmlstep)
```

```{r}
plot(lmlstep,which=1)
```
```{r}
plot(lmlstep,which=2)
```
```{r}
plot(lmlstep,which=3)
```
```{r}
plot(lmlstep,which=5)
```
## Analysis with Centered data

Scale all of the variables except Y.  Our Y is the last variable

```{r}
UScrimeCC <- scale(UScrime[,-ncol(UScrime)])
## Add crime back as unscaled.
UScrimeC <- as.data.frame(cbind(UScrimeCC,Crime=UScrime[,ncol(UScrime)]))
summary(UScrimeC)
```

```{r}
lmlstepC <- lm(formula(lmlstep),data=UScrimeC)
summary(lmlstepC)
```

## Write out data for SPSS analysis

Want to add the centered data as it is faster to do that here than in SPSS.
Change the names on `UScrimeC` to a common pattern, adding "C" to the end of each.

```{r}
colnames(UScrimeCC) <- paste(colnames(UScrimeCC),"C",sep="")
haven::write_sav(cbind(UScrime,UScrimeCC),"UScrime.sav")
```

