---
title: "Diabetes Risk"
author: "Russell Almond"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Diabetes Risk Data

These data come from the [UC Irvine Machine Learning Repository](https://archive.ics.uci.edu/dataset/529/early+stage+diabetes+risk+prediction+dataset). 
There are related to patients in a certain diabetes hospital in Bangledesh.  It is not
explicitly stated, but it probably refers to Type II diabetes (as Type I is usually
identified in childhood).

We will look at the risk of diabetes using three variables:  `Age`, `Sex` (the 
phenotype not the gender expression) and `Obesity`.  

## Read and clean data

```{r}
read_csv("diabetes_data_upload.csv") |> 
  mutate(Diabetic=as.factor(class),Sex=factor(Gender),Obese=factor(Obesity)) |>
  select(all_of(c("Age","Sex","Obese","Diabetic"))) ->
  DData
summary(DData)
```

## Exploratory Analyses

```{r}
ggplot(DData,aes(x=Age,y=as.numeric(Diabetic))) +
  geom_point(position=position_jitter(width=.5,height = .1)) +
  geom_smooth()
```

```{r}
ggplot(DData,aes(x=Age,y=as.numeric(Diabetic),color=Sex)) +
  geom_point(position=position_jitter(width=.5,height = .1)) +
  geom_smooth()
```

```{r}
ggplot(DData,aes(x=Age,y=as.numeric(Diabetic),color=Obese)) +
  geom_point(position=position_jitter(width=.5,height = .1)) +
  geom_smooth()
```

Only two different variables, so we can get them all on the same plot:  (If it is not clear from the legend, the dashed line is `Obese==Yes`.

```{r}
ggplot(DData,aes(x=Age,y=as.numeric(Diabetic),color=Sex,shape=Obese,linetype=Obese)) +
  geom_point(position=position_jitter(width=.5,height = .1)) +
  geom_smooth()
```

# First model

```{r}
glmAge <- glm(Diabetic~Age,data=DData,family=binomial())
summary(glmAge)
```
I'm going to do this a lot, so might as well write a function.
```{r}
APAsum <- function (model) {
  df <- model$df.null-model$df.residual
  X2 <- model$null.deviance- model$deviance
  p <- 1-pchisq(X2,df)
  paste("X^2(",df,") =", round(X2,2), 
        ifelse(p<.001, ", p < .001", paste(", p =",round(p,3))))
}
cat(APAsum(glmAge),"\n")
```

# Now Add Sex

I'll directly fit the interaction model and maybe simplify it later.

```{r}
glmAgeSex <- glm(Diabetic~Age*Sex,data=DData,family=binomial())
summary(glmAgeSex)
cat(APAsum(glmAgeSex),"\n")
```

## Same trick with Obesity

```{r}
glmAgeFat <- glm(Diabetic~Age*Obese,data=DData,family=binomial())
summary(glmAgeFat)
cat(APAsum(glmAgeFat),"\n")
```

## Model Search

We will consider the full model including three-way interactions:

```{r}
glmStep <- step(glmAge,list(lower=Diabetes~1,  # constant risk
                            upper=Diabetes~Age*Sex*Obese),
                trace=2)
```

## Final model

```{r}
summary(glmStep)
cat(APAsum(glmStep),"\n")
```

## Predictions

```{r}
risktab <- data.frame(Age=rep(15+5*(1:11),each=4),
                  Sex=factor(rep(rep(levels(DData$Sex),each=2),11)),
                  Obese=factor(rep(levels(DData$Obese),22)))
head(risktab)
```


```{r}
risktab$Risk <- psych::logistic(predict(glmStep,risktab))
head(risktab)
```

This will be easier to look at if we turn `Sex` and `Obese` into columns.  Can do this with the pivot function.

In the chart below "Yes" and "No" refer to Obsese, so "Male_Yes" is an obese male, 
and "Male_No" is a non-obese male.

```{r}
riskchart <- pivot_wider(risktab,id_cols=Age,names_from=all_of(c("Sex","Obese")),
                         values_from=Risk)
knitr::kable(riskchart,digits=2)
```


Save the data out for use in SPSS.

```{r}
haven::write_sav(DData,"diabetesRisk.sav")
```


