---
title: "EDF 5401 Midterm, Part 1: Hurricanes."
date: 2013-10-15
format: pdf
---

These data come from the [Data and Story Library](https://dasl.datadescription.com/datafile/hurricanes-2015/?_sf_s=hurricane&_sfm_cases=4+59943)

Most weather models note at relationship between the barimetric pressure and
the peak wind speeds.  A secondary question is, as the average temperature
rises, is that relationship changing.

```{r}
library(tidyverse)
library(DescTools)
```


# Part 1 Hurricanes

## Exploratory Analysis

### Load the data
Load the data.  Force category to be an ordered category.
```{r hload}
hurric <- read_delim("hurricanes-2015.txt")
hurric$Category <- ordered(hurric$Category)
summary(hurric)
```

### One-dimensional analyses

```{r descPress}
Desc(hurric$`Central.Pressure(mb)`)
```

Look at differences in pressure by category.

```{r pressCat}
ggplot(hurric, aes(x=`Central.Pressure(mb)`,y=Category)) + geom_violin()
```

```{r descSpeed}
Desc(hurric$`Max.Wind.Speed(kts)`)
```

Look at differences in maximum speed by category.  (Note category is largely
defined by wind speed.)

```{r speedCat}
ggplot(hurric, aes(x=`Max.Wind.Speed(kts)`,y=Category)) + geom_violin()
```

### Relationships with time

```{r hurrCor}
round(cor(hurric[,2:4],use="complete.obs"),3)
```

```{r speedYear}
ggplot(hurric,aes(x=Year,y=`Max.Wind.Speed(kts)`)) + 
  geom_point() + geom_smooth()
```

Hmm.  Note big gap in data between 1925 and 1975.  Maybe before/after climate
change?  Note 1950 appears to be a cut point.

```{r recent}
hurric <- mutate(hurric,recent=Year>1950)
```


```{r pressYear}
ggplot(hurric,aes(x=Year,y=`Central.Pressure(mb)`)) + 
  geom_point() + geom_smooth()
```

## Scatterplots

### XY

```{r hurscat}
ggplot(hurric,aes(x=`Central.Pressure(mb)`,y=`Max.Wind.Speed(kts)`)) +
  geom_point() + geom_smooth()
```

### Jittered

Hmm.  Points lying on top of each other, try some jittering.

```{r hurscat1}
ggplot(hurric,aes(x=`Central.Pressure(mb)`,y=`Max.Wind.Speed(kts)`)) +
  geom_point(position="jitter") + geom_smooth()
```
### XY by Recent

Color by recent to see if the current and recent groups are similar or not.

```{r hurscat2}
ggplot(hurric,aes(x=`Central.Pressure(mb)`,y=`Max.Wind.Speed(kts)`,color=recent)) +
  geom_point(position="jitter") + geom_smooth()
```
### Outliers

There seem to be a couple of ouliers.  Lets try to find them.

```{r hurricOut}
hout <- which(hurric$`Central.Pressure(mb)`<960 & 
                hurric$`Max.Wind.Speed(kts)` < 75)
hurric[hout,]
```
```{r hurscat2out}
ggplot(hurric,aes(x=`Central.Pressure(mb)`,y=`Max.Wind.Speed(kts)`,
                  color=recent)) +
  geom_point(position="jitter") + geom_smooth() +
  geom_text(data=hurric[hout,],aes(label=hout))
```

## Build the Regression Model

```{r hurriclm}
lm_hurric <- lm(`Max.Wind.Speed(kts)` ~ `Central.Pressure(mb)`, data=hurric)
summary(lm_hurric)
```

```{r diagHurric}
oldpar <- par(mfrow=c(2,2))
plot(lm_hurric)
par(oldpar)
```

### Try some Y transforms

Log
```{r logHurricScat}
ggplot(hurric,aes(x=`Central.Pressure(mb)`,y=`Max.Wind.Speed(kts)`,color=recent)) +
  geom_point(position="jitter") + geom_smooth() + scale_y_log10()
```

Log model
```{r loglmHurric}
llm_hurric <- lm(log(`Max.Wind.Speed(kts)`,10) ~ `Central.Pressure(mb)`, data=hurric)
summary(llm_hurric)
```
Log Diagnostics

```{r diaglogHurric}
oldpar <- par(mfrow=c(2,2))
plot(llm_hurric)
par(oldpar)
```

Flipping a coin, I'm doing the rest of the analyses on the linear scale.

### Outliers

Calculate dfbetas for identified outliers
```{r dfbetas}
dfbetas(lm_hurric)[hout,]
```
Run the regression without the outliers.
```{r lmHurricNO}
lm_hurric_no <-  lm(`Max.Wind.Speed(kts)` ~ `Central.Pressure(mb)`, 
                    data=hurric, subset=-hout)
summary(lm_hurric_no)
```

## Run separately for old and recent data.

Redo the plot with `method="lm"` to visualize different lines.
```{r diffPlot}
ggplot(hurric,aes(x=`Central.Pressure(mb)`,y=`Max.Wind.Speed(kts)`,color=recent)) +
  geom_point(position="jitter") + geom_smooth(method="lm")
```

### 19th C, Early 20th

```{r lmEarly}
lm_hurric_19 <-  lm(`Max.Wind.Speed(kts)` ~ `Central.Pressure(mb)`, 
                    data=hurric, subset=!recent)
summary(lm_hurric_19)
## Save slope and SE for later processing.
hurric_slope_19 <- summary(lm_hurric_19)$coefficients[2,1:2]
```

### Late 20th, Early 21st
```{r lmLate}
lm_hurric_20 <-  lm(`Max.Wind.Speed(kts)` ~ `Central.Pressure(mb)`, 
                    data=hurric, subset=recent)
summary(lm_hurric_20)
## Save slope and SE for later processing.
hurric_slope_20 <- summary(lm_hurric_20)$coefficients[2,1:2]
```

Compare slopes in a table:
```{r slopetab}
rbind(early=hurric_slope_19,
      late=hurric_slope_20)
```
Standard error for the difference is $\sqrt{s_1^2 +s_2^2}$

```{r sediff}
sqrt(hurric_slope_19[2]^2+hurric_slope_20[2]^2)
```


