---
title: "Midterm Notes"
format: html
---

## Digits

Statistics shouldn't have a lot more significant digits than the the raw data.

Means (medians, slopes & intercept)

:   One more digit than the raw data

Standard Deviations (standard errors)

:   Two more digits

```{r}
summary(state.x77)
```

```{r}
mean(state.x77[-2,"Income"])
round(mean(state.x77[-2,"Income"]),1)
```

```{r}
sd(state.x77[-2,"Income"])
round(sd(state.x77[-2,"Income"]),2)
```

### Scientific notation

```{r}
cat("p=",sqrt(.Machine$double.eps),"\n")
```

If the p-value goes into scientific notation, write $p < .001$

## APA Style

Expressions in \[math\] italics.

```{r}
summary(lm(Income~Frost,data=as.data.frame(state.x77)))
```

### Effect sizes

For regression, the effect sizes are $r$ or $R^2$ (or partial $r$).

### t-test

Template: ($t(\text{df})=\text{t}, p=\text{p}$) or $p<.001$

df is (residual) degrees of freedom

p is p-value.

$$
t(48) = 1.609, p=0.114
$$

### F-test

Template: ($F(\text{dfn},\text{dfd})=\text{F}, p=\text{p}$)

dfn is degrees of freedom in numerator (usually number of predictors).

dfd is the residual degrees of freedom

Multiple R-squared: 0.0512, Adjusted R-squared: 0.03144 F-statistic: 2.59 on 1 and 48 DF, p-value: 0.1141

$$ F(1,48) = 2.59, R^2=0.051, p=.114 $$

### Correlation

## Logs

### Log in the equation

$$ \hat Y = b_0 + b_1\log(X)$$

$$ \widehat{\log(Y)} = b_0 + b_1 X $$

### Log10 vs natural log

In R, `log()` natural log, `log10()` is common log.

In SPSS, `log` is common log (I think), and `ln` is natural log.

Difference is a constant (`log(10)` = `{r} log(10)`)

Common logs are generally easier to interpret: change of $1$ on common log scale is a factor of 10.

In the context of linear regression in the social sciences, Gelman and Hill write\[1\]:

> We prefer natural logs (that is, logarithms base ğ‘’) \[for $\log(Y)$\] because, as described above, coefficients on the natural-log scale are directly interpretable as approximate proportional differences: with a coefficient of 0.06, a difference of 1 in ğ‘¥ corresponds to an approximate 6% difference in ğ‘¦, and so forth.

\[1\] Andrew Gelman and Jennifer Hill (2007). Data Analysis using Regression and Multilevel/Hierarchical Models. Cambridge University Press: Cambridge; New York, pp. 60-61. \[hat tip `fmark@StackExchange`\]

When in doubt, plug some numbers into the regression equation.
