---
title: "EDF 5401 Midterm, Part 2: Depression and the internet"
date: 2013-10-15
format: pdf
---

These data come from the [Data and Story Library](https://dasl.datadescription.com/datafile/depression-and-the-internet/?_sf_s=depression&_sfm_cases=4+59943)

These household were given free internet access in return for agreeing to being 
tracked.  Several statistics were reported before and after.

Note that in 1998, home internet access was probably dial-up and not the
cable/DSL speeds of the current internet.

```{r}
library(tidyverse)
library(DescTools)
```

# Part 2: Depression and the Internet

## Analysis

### Load the data
Load the data.  Force category to be an ordered category.
```{r dload}
depress <- read_delim("depression-and-the-internet.txt")
depress$Gender <- factor(depress$Gender)
depress$Age <- factor(depress$Age)
depress$Race <- factor(depress$`Race (white = 1, minority = 0)`,0:1,
                       c("minority","white"))
summary(depress)
```

### One Dimensional Summaries

```{r dInternet}
Desc(depress$`Internet use (mean hours per week)`)
```


```{r dBefore}
Desc(depress$DepressionBefore)
```

Identify outliers
```{r hbefore}
hbefore <- which (depress$DepressionBefore > 2)
hbefore
```


```{r dAfter}
Desc(depress$DepressionAfter)
```

Identify outliers
```{r hafter}
hafter <- which (depress$DepressionAfter > 1.5)
hafter
```

Subtract before from after to get change.
```{r dChange}
depress$DepressionChange <- depress$DepressionAfter - depress$DepressionBefore
Desc(depress$DepressionChange)
```
Identify outliers  (There is one point near the lower fence at -1.13, 
mark anything beyond that as an outlier.)
```{r hchange}
hchange <- which (abs(depress$DepressionChange) > 1.25)
hchange
```

### Relationships with Age

```{r aInternet}
ggplot(depress,aes(x=`Internet use (mean hours per week)`,y=Age)) +
  geom_violin()
```

```{r alInternet}
ggplot(depress,aes(x=`Internet use (mean hours per week)`,y=Age)) +
  geom_violin() + scale_x_log10()
```

```{r aBefore}
ggplot(depress,aes(x=DepressionBefore,y=Age)) +
  geom_violin()
```
```{r aAfter}
ggplot(depress,aes(x=DepressionAfter,y=Age)) +
  geom_violin()
```

```{r aChange}
ggplot(depress,aes(x=DepressionChange,y=Age)) +
  geom_violin()
```

## Scatterplots

### X, Y

```{r dxy}
ggplot(depress,aes(x=`Internet use (mean hours per week)`,
                   y=DepressionChange)) +
  geom_point() + geom_smooth()+
  geom_text(data=depress[hchange,],aes(label=hchange))
```


### log(X), Y

```{r dlogxy}
ggplot(depress,aes(x=`Internet use (mean hours per week)`,
                   y=DepressionChange)) +
  geom_point() + geom_smooth()+ scale_x_log10() +
  geom_text(data=depress[hchange,],aes(label=hchange))
```

Hmm.  Note a number of points piled up on the $y$-axis.  The problem is 
$\log(0) = -\infty$.  The solution is to add a small number before taking the
logs.  Looks like most values are above .001, so add .0001.

An alternative would be to exclude these people from the model.

```{r dlogx1y}
ggplot(depress,aes(x=`Internet use (mean hours per week)`+.0001,
                   y=DepressionChange)) +
  geom_point() + geom_smooth()+ scale_x_log10() +
  geom_text(data=depress[hchange,],aes(label=hchange))
```

### By Age

```{r dxyAge}
#| layout-ncol: 2
ggplot(depress,aes(x=`Internet use (mean hours per week)`,
                   y=DepressionChange, color=Age)) +
  geom_point() + geom_smooth() +
  geom_text(data=depress[hchange,],aes(label=hchange))
ggplot(depress,aes(x=`Internet use (mean hours per week)`+.0001,
                   y=DepressionChange,color=Age)) +
  geom_point() + geom_smooth()+ scale_x_log10() +
  geom_text(data=depress[hchange,],aes(label=hchange))
```


### By Gender

```{r dxyGender}
#| layout-ncol: 2
ggplot(depress,aes(x=`Internet use (mean hours per week)`,
                   y=DepressionChange, color=Gender)) +
  geom_point() + geom_smooth() +
  geom_text(data=depress[hchange,],aes(label=hchange))
ggplot(depress,aes(x=`Internet use (mean hours per week)`+.0001,
                   y=DepressionChange,color=Gender)) +
  geom_point() + geom_smooth()+ scale_x_log10() +
  geom_text(data=depress[hchange,],aes(label=hchange))
```

### By Race (Whiteness)

Race is coded white/non-white.  So this is essentially looking at the effect of 
"whiteness" (whatever that is).

```{r dxyRace}
#| layout-ncol: 2
ggplot(depress,aes(x=`Internet use (mean hours per week)`,
                   y=DepressionChange, color=Race)) +
  geom_point() + geom_smooth() +
  geom_text(data=depress[hchange,],aes(label=hchange))
ggplot(depress,aes(x=`Internet use (mean hours per week)`+.0001,
                   y=DepressionChange,color=Race)) +
  geom_point() + geom_smooth()+ scale_x_log10() +
  geom_text(data=depress[hchange,],aes(label=hchange))
```
## Regression Model

```{r dlm}
dlm <- lm(DepressionChange ~ `Internet use (mean hours per week)`,
          data=depress)
summary(dlm)
```

```{r dlmplots}
#| layout-nrow: 2
#| layout-ncol: 2
plot(dlm)
```


### Log X

```{r dllm}
dllm <- lm(DepressionChange ~ log(`Internet use (mean hours per week)`+.0001),
          data=depress)
summary(dllm)
```


```{r dllmplots}
#| layout-nrow: 2
#| layout-ncol: 2
plot(dllm)
```
### Leverage Points

Linear scale

```{r dfblm}
dfblm <- dfbetas(dlm)
hlev <- which(abs(dfblm[,2])>2/sqrt(nrow(dfblm)))
dfblm[hlev,]
```
Rerun without leverage points.

Linear model.
```{r dlmnl}
summary(lm(DepressionChange ~ `Internet use (mean hours per week)`,
          data=depress,subset=-hlev))
```

```{r dllmnl}
summary(lm(DepressionChange ~ log(`Internet use (mean hours per week)`+.0001),
          data=depress,subset=-hlev))
```

## Sensitivity Analyses

### Outliers

Look at sensitivity to the outliers identified above.

```{r dlmno}
summary(lm(DepressionChange ~ `Internet use (mean hours per week)`,
          data=depress,subset=-hchange))
```


```{r dllmno}
summary(lm(DepressionChange ~ log(`Internet use (mean hours per week)`+.0001),
          data=depress,subset=-hchange))
```
### Internet Vegans

Take a closer look at the people who spend 0 hours per week on the internet.
Call them _Internet Vegans_.  Look at the sensitivity to them.

```{r dlmnv}
vegans <- which(depress$`Internet use (mean hours per week)`<.0005)
summary(lm(DepressionChange ~ `Internet use (mean hours per week)`,
          data=depress,subset=-vegans))
```
```{r dllmnv}
summary(lm(DepressionChange ~ log(`Internet use (mean hours per week)`+.0001),
          data=depress,subset=-vegans))
```
