---
title: "EDF 5401 Practice Final"
author: "Russell Almond"
date: "`r Sys.Date()`"
output: word_document
---

# Physics Playground Background

## The Game

_Physics Playground_ is a game for teaching physics.  One part of the program is 
learning supports.  There were two kinds of learning supports:

* _Cogntive Supports_:  Short videos explaining the physics.
* _Affective Supports_:  Motivational messages and diversions designed to help 
alleviate frustration.

There are matched pretests and posttests which are divided into four parts:

* `NearECT` --- Energy can transfer (ECT), near transfer (similar to the game) items
* `FarECT` --- Energy can transfer (ECT), far transfer (not similar to the game) items
* `NearPOT` --- Properties of torque (POT), near transfer
* `FarPOT` --- Properties of torque (POT), far transfer

More information about the game and the experiment can be found at 
<https://pluto.coe.fsu.edu/PhysicsDataPlayground/>.

## The support efficacy experiment and key variables

The experiment is a bit complex, with half of the student getting the ECT game levels 
and posttest in the first half of the experiment and the others in the second half.  
For simplicity we will work with just the ECT-first:  `Topic1=="ECT"`.

For these students, there are three condition groups (`Condition`):

* `A` -- Cognitive Supports Only
* `C` -- Cognitive and Affective Support
* `F` -- No supports (control)

The variables, `CognitiveSupports` and `AffectiveECT` are logical variables describing
whether or not the cognitive supports and the affective supports were present (during 
the first half of the study).

The pretest score is labeled `ECT` and the posttest score is labeled `ECTpost`.  
There is also a `NearECT` and `NearECTpost` (just the near transfer items) and
a `FarECT` and `FarECTpost` (far transfer items).

## Additional covariates

`Age`
: Self-reported age in years.

`Sex`, `Gender`
: Self-reported gender

`White`
: Whether or not the subject identified as White.

`Gaming`
: "How often do you play video games?"  Ordinal, possible values:   
`Never`,	`Once a month or less`,	`Once a week`,	`3-4 times a week`,	`Every day`

`Physics`
: "Have you studied Physics?" Possible answers `Yes`, `No`.

`POT`
: Score on the properties of torque physics pretest.

`PhysicsScore`
: `POT + ECT`

`IMI_Enj`
: Enjoyment subscale of the intrinsic motivation inventory (IMI)

`IMI_Eng`
: Engagement subscale of the IMI

`IMI_PC`
: Player confidence subscale of the IMI

`IMI_Effort` 
: Effort subscale of the IMI

`IMI_Frust`
: Frustration subscale of the IMI

`IMI_Value`
: Value of experience subscale IMI

`gold`, `silver`, `quit`
: The number of game levels in which the player received a gold trophy 
(solved the level with an efficient solution), silver trophy (solved the level, but 
did not meet the qualifications for efficiency), or abandoned the level without
solving it.

## Data Loading and cleaning

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(DescTools)
library(rgl)
library(GGally)
```

Load the data (from the internet)

```{r load}
source("https://pluto.coe.fsu.edu/PhysicsDataPlayground/Data/AllMetadata.R")
PPIESFall2022 <- 
  read_csv("https://pluto.coe.fsu.edu/PhysicsDataPlayground/Data/PPIESFall2022Full.csv",
           col_types=colAll)
```

Extract a subset of the data with just the cases/variables we are working with.

```{r}
PPIESExtract <- PPIESFall2022 |>
  filter(Topic1=="ECT") |>
  select(all_of(c("StudyID", "Condition", "CognitiveSupports", "AffectiveECT",
                  "Age", "Sex", "White", "Gaming", "Physics", 
                  "NearECT", "FarECT", "ECT",  "NearECTpost", "FarECTpost",
                  "ECTpost", "POT", "PhysicsScore",
                  "IMI_Enj", "IMI_Eng", "IMI_PC", "IMI_Effort", "IMI_Frust",
                  "IMI_Value",
                  "gold", "silver", "quit"))) |>
  na.omit()
summary(PPIESExtract)         
```

Only three students who identify as nonbinary.  Too small to analyze, so take them 
out.

## Derieved variables

Gain Scores
```{r gains}
PPIESExtract <- PPIESExtract |>
  mutate(ECTgain=ECTpost-ECT,
         NearECTgain=NearECTpost-NearECT,
         FarECTgain=FarECTpost-FarECT,
         LowECT= ECT<median(ECT))
```

The reason for LowECT will become apparent later.

```{r CleanGender}
PPIESExtract <- filter(PPIESExtract,Sex %in% c("Male","Female"))
```

### Some changes for interpretability

Ages run from 12-16 (middle and high school students).   Create an new variable 
`Age1` for subtracts 0, so it is essential years of schooling since elementary 
school.

Center physics variable to aid in interpretation.

```{r Age1}
PPIESExtract <- mutate(PPIESExtract,Age1=Age-12,
                       PhysicsCenter=PhysicsScore-mean(PhysicsScore))
```

It will be convenient (mainly for graphics) to have better labeled condition variable.

```{r supports}
PPIESExtract$Supports <- 
  factor(PPIESExtract$CognitiveSupports+PPIESExtract$AffectiveECT,
         levels=0:2,labels=c("None","COG only","COG+AFF"))
```

## Enjoyment definition

For the logistic regression, we will define Enjoyed as `IMI_Enj >7` (This ranges from
2--14).

```{r Enj}
PPIESExtract <- mutate(PPIESExtract,Enjoyed=IMI_Enj>8)
```


Write this out in SPSS format so we can do a parallel analysis there.
```{r}
haven::write_sav(PPIESExtract,"PPIESExtract.sav")
```



# ANCOVA

The primary analysis will be an ANCOVA with `ECTpost` as the dependent variable, 
`ECT` as the covariate, and `CognitiveSupport` and `AffectiveECT` as the treatment 
variables.  In addition we will explore other possible covariates.

## Exploratory Analysis

```{r ECTscatter}
ggplot(PPIESExtract,aes(x=ECT,y=ECTpost)) + geom_point(position="jitter") +
  geom_smooth()
```

Treatment by condition breakdown

```{r ECTprepostbyCond}
ggplot(PPIESExtract,aes(x=ECT,y=ECTpost,color=Supports)) +
  geom_point(position="jitter") +
  geom_smooth()

```


### Demographic Breakdowns

```{r ECTbyGender}
ggplot(PPIESExtract,aes(x=ECT,y=ECTpost,color=Sex)) + 
  geom_point(position="jitter") +
  geom_smooth()
```

```{r ECTbyAge}
ggplot(PPIESExtract,aes(x=ECT,y=ECTpost,color=factor(Age))) + 
  geom_point(position="jitter") +
  geom_smooth()
```
```{r ECTbyRace}
ggplot(PPIESExtract,aes(x=ECT,y=ECTpost,color=White)) + 
  geom_point(position="jitter") +
  geom_smooth()
```

```{r ECTbyGaming}
ggplot(PPIESExtract,aes(x=ECT,y=ECTpost,color=Gaming)) + 
  geom_point(position="jitter") +
  geom_smooth()
```
```{r ECTbyPhysics}
ggplot(PPIESExtract,aes(x=ECT,y=ECTpost,color=Physics)) + 
  geom_point(position="jitter") +
  geom_smooth()
```

Look at the correlation between the trophy variables and ECT with a scatterplot 
matrix.

```{r ECTbyTropies}
#| warning: false
#| message: false
select(PPIESExtract,all_of(c("ECTpost","ECT","gold","silver","quit"))) |>
  ggpairs()
```

### Treatment Effect 

A = Cognitive Only
C = Cogntivie + Affective
F = Control (no supports)

```{r}
ggplot(PPIESExtract,aes(x=Supports,y=ECTgain)) + geom_boxplot()

```
## Baseline Model

```{r}
ECTbase <- lm(ECTpost ~ ECT + Supports,data=PPIESExtract)
summary(ECTbase)
```

## ATI Model

```{r ATI}
ECTATI <- lm(ECTpost ~ ECT * Supports,data=PPIESExtract)
summary(ECTATI)
```
```{r ATIAnova}
anova(ECTbase,ECTATI)
```
## Diagnostic plots

```{r plot.lm}
oldpar <- par(mfrow=c(2,2)) # Put all of the plot together
plot(ECTbase)
par(oldpar) # Restore the graphics window

```
## Influential points

```{r dfbetas}
dfb <- dfbetas(ECTbase)
summary(dfb)
```

Flagging value is `{r} 2/sqrt(nrow(dfb))`

```{r flagging}
## Calculate a logical value rows higer than flagged value.
flags <- apply(dfb,1,function(r) any(r>2/sqrt(nrow(dfb))))
PPIESExtract$StudyID[flags]
dfb[flags,]
```


### Remove flagged values

```{r noinf}
ECTnoinf <- lm(ECTpost ~ ECT + Supports, data=PPIESExtract,
               subset = !flags)
summary(ECTnoinf)
```

```{r coefficent table}
rbind(base=coef(ECTbase),
      noinf=coef(ECTnoinf))
```

## Exploring more models

```{r search}
ECTminmod <- ECTpost ~ ECT + Supports
ECTmaxmod <- ECTpost ~ ECT + Supports + Age + Sex + White + Gaming +
  Physics + POT + gold + silver + quit
ECTstep <- step(ECTbase,list(lower=ECTminmod,upper=ECTmaxmod),trace=2)
```

```{r}
summary(ECTstep)
```

## Splitting into High and Low initial ECT values

Added the median split variable so that we could fit separate slopes to high and 
low ECT values.

```{r highlow}
ECThighlow <- lm(ECTpost ~ ECT*LowECT + Supports*LowECT,data=PPIESExtract)
summary(ECThighlow)
```

```{r lowAnova}
anova(ECTbase,ECThighlow)
```

# Logistic Regression

The research question is what is the relationship between game enjoyment and physics 
ability (as measured by the pretest), `PhysicsScore`

The `IMI_Enj` variables is the sum of two 7-option Likert scale items, so 8 is 
neutral on both variables.  So define enjoyment as `IMI_Enj > 8`.

Also, look at the role of gender.

## Exploratory analysis

```{r EnjoyXPhysics}
ggplot(PPIESExtract,aes(x=PhysicsScore,y=as.numeric(Enjoyed))) +
  geom_point(position=position_jitter(width=.5,height=.1)) +
  geom_smooth()
```

Enjoyment by Gender

```{r EnjGender}
table(PPIESExtract$Enjoyed,PPIESExtract$Sex)
```

All three variables together

```{r EnjPhySex}
ggplot(PPIESExtract,aes(x=PhysicsScore,y=as.numeric(Enjoyed),color=Sex)) +
  geom_point(position=position_jitter(width=.5,height=.1)) +
  geom_smooth()
```

## Base Model

```{r logPhy}
ENJphys <- glm(Enjoyed ~ PhysicsCenter, data=PPIESExtract,
              family=binomial())
summary(ENJphys)
```

Prediction at mean physics ability
```{r pred}
round(psych::logistic(coef(ENJphys)[1]),2)
```


## Model with gender

```{r logPhyGen}
ENJphysG <- glm(Enjoyed ~ PhysicsCenter + Sex, data=PPIESExtract,
              family=binomial())
summary(ENJphysG)
```
Prediction at mean physics ability

```{r predG}
round(psych::logistic(coef(ENJphysG)[1]+c(Male=0,Female=1)*coef(ENJphysG)[3]),2)
```

## Model with gender interaction

```{r logPhyXGen}
ENJphysXG <- glm(Enjoyed ~ PhysicsCenter * Sex, data=PPIESExtract,
              family=binomial())
summary(ENJphysXG)
```
Prediction at mean physics ability

```{r predXG}
round(psych::logistic(coef(ENJphysXG)[1]+c(Male=0,Female=1)*coef(ENJphysXG)[3]),2)
```


## Analysis of Deviance

Like Analysis of Variance, only with Deviance & chi-squared.

```{r AOD}
anova(ENJphys,ENJphysG,ENJphysXG)
```

The differences in deviance are chi-squared valued.

Here is a chi-square table.
```{r chi2}
chi2.1.table <- rbind(prob=c(.5,.75,.9,.95,.99),
                      X2=qchisq(c(.5,.75,.9,.95,.99),1))
round(chi2.1.table,2)

```



