---
title: "ACED Simple Regression"
format: html
---
```{r library}
library(DescTools)
library(tidyverse)
```

# ACED Data
```{r loadACEDnoNA}
ACEDextract <- read_csv("ACED_extract1.csv",na="-999")
ACEDextract$Session <- factor(ACEDextract$Session)
ACEDextract$Cond_code <- factor(ACEDextract$Cond_code)
ACEDextract$Sequencing <- factor(ACEDextract$Sequencing)
ACEDextract$Feedback <- factor(ACEDextract$Feedback)
ACEDextract$Gender <- factor(ACEDextract$Gender)
ACEDextract$Race <- factor(ACEDextract$Race,1:8)
ACEDextract$Level_Code <- factor(ACEDextract$Level_Code)
```

```{r mutate}
ACEDextract %>%
  mutate(gain=post_scaled-pre_scaled) ->
  ACEDextract
```

## Research Questions

# Making Scatterplots

## Scatterplot

## Adding lines and smooths

## Jittering

## Coloring points

# Fitting a Linear Model

## Formulas

## Other arguments of `lm()`

* `data`

* `subset`

* `weights`

* `na.action`

## components

* `coefficients`

* `residuals`

* `fitted.values`

* `df.residuals`

## Summaries

* `print.lm`

* `summary.lm`

* `anova.lm`

* `plot.lm`

## Extracting bits

* `coef`

* `effects`

* `vcov`

* `nobs`

* `variable.names`

## Extracting bits from the summary

* `summary()$sigma` -- residual sd/standard error of the estimate

* `summary()$df` -- degrees of freedom

* `summary()$fstatistic`

* `summary()$r.squared`, `summary()$adj.r.squared`

## Prediction

* `predict`

* `fitted`

* `residuals`, `rstandard`, `rstudent`

* `simulate`

## Diagnostics

* `dfbeta`, `dfbetas`, `dffits`

* `cooks.distance`

* `influence`

* `hatvalues`

## Model Fit

* `logLik`

* `deviance`

# Tasks

1. Make marginal summaries for the following variables:

Correct,Incorrect,Reamaining,ElapsedTime,pre_scaled,post_scaled,EAP.sgp

2. Same as above, but break down by Level_code

