---
title: "Case Study 6, Logistic Regression"
format: html
---

# Generalized Linear Models

## GLM families

* _Distribution_ -- A probability distribution $P(\boldsymbol{Y}|\boldsymbol{\mu},\boldsymbol{\nu})$
  - $\boldsymbol{\mu} = E[\boldsymbol{Y}|\textbf{X}]$ -- expected value
  - $\tau$ is a scale parameter (e.g., variance)
* _Linear Predictor_ -- $\boldsymbol{\eta} = \textbf{X}\boldsymbol{\beta}$
* _Link Function_ -- $g(\boldsymbol{\mu}) = \bolsymbol{\eta}$, or 
$\boldsymbol{\mu} = g^{1}(\boldsymbol{\mu})$
* _Variance Function_ -- $\boldsymbol{\nu} = V(\boldsymbol{\mu},\tau)$, where $\tau$ is a parameter estimated from the data.

## Example:  Logistic Regressions

* _Distribution_:  Binomial distribution, with parameter $p$.  Expected value is $p$ and variance is $p(1-p)$.

* _Linear Predictor_:  $\boldsymbol{\eta} = \textbf{X}\boldsymbol{\beta}$

* _Link Function_: Logit (log odds) function.  $\boldsymbol{p} = \text{logit}^{-1}(\boldsymbol{\eta})$.

$$ \text{logit}(p) = \ln \frac{p}{1-p}$$

```{r}
curve(psych::logit(x),xlim=c(0,1))
```

Its inverse is the _logistic_ function

$$ \text{logit}^{-1}(x) = \frac{1}{1+\exp(-x)}$$
```{r}
curve(psych::logistic(x),xlim=c(-6,6))
```

* _Variance_ is $p(1-p)$


## Families

* `gaussian(link="identity")` -- Ordinary Regression
* `binomial(link="logit")` -- Logistic Regression:  True/False
* `binomial(link="probit")` -- Probit Regression:  True/False
* `poisson(link="log")` -- Loglinear models:  Contingency Tables


# Heart Data
This example is based on a study by @janosi1989 and recorded in the UCI Machine Learning Repository [@murphy1992]. The direct link is: <https://archive.ics.uci.edu/dataset/45/heart+disease>
```{r Rpacks}
library(tidyverse)
```

Read in the Cleveland data set from @janosi1989. The `heart.features` is information about the coding from the web site.

```{r DownloadData}

heart.features <- list(
  Age=numeric(),
  Sex=c(Male=1,Female=0),
  CP=c(TypicalAnginal=1, AtypicalAnginal=2, NonAnginal=3, Asymptomatic=4),
  trestbps=numeric(),
  chol=numeric(),
  fbs=c(high=1,normal=0),
  restecg=c(Normal=0, STTabnormality=1, LeftVentricularHypertrophy=2),
  thalach=numeric(),
  exang=c(Yes=1, No=0),
  oldpeak=numeric(),
  slope=c(down=1, flat=2, up=3),
  ca=c(Zero=0, One=1, Two=2, Three=3, Four=4),
  thal=c(Normal=3, FixedDefect=6, ReversibleDefect=7),
  health=c(Healthy=0, S1=1, S2=2, S3=3, S4=4)
)

cleveland <- read_csv("processed.cleveland.csv",
                      col_names=names(heart.features),
                      na="?")
cleveland
```

```{r recode}
for (f in names(heart.features)) {
  if (length(heart.features[[f]])>0L) {
    cleveland[[f]] <- factor(cleveland[[f]],heart.features[[f]],
                             names(heart.features[[f]]))
  }
}

cleveland
```


